Quantifying the rock-paper-scissors effect

### 2025-01-30 inst/rock_paper_scissors_monster.Rmd
consistentify "time to process" information
more explanation

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about an hour to run without cache.  Here I use the `hyper2`
package to quantify the strength of a circular rock-paper-scissors
effect using the `hyper2` package, improving on previous results
[@west2008] which merely established its existence.  I use two
datasets, a three-player scoreline and a synthetic dataset referring
to wear on a seven-tooth gear wheel.

## Rock Paper Scissors

@@

## RPS in chess

@@

Estimating $p_M$ is also straightforward:

$$
Prob(\mathrm{procyclic}\ \mathrm{win})=\frac{\alpha+M}{2\alpha+M}
$$

(a "procyclic win" is a win that would be expected under a strong
monster).  Here we also have $3\alpha+M=1$, so to estimate the
strength of the monster we would solve

@@

### Profile likelihood

@@

### Analysis of seven-tooth gear with `hyper3`

>>>>

<font size="1"> (takes about an hour to run without cache)
</font>

### Abstract {-}

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  Here I use the `hyper2` package to quantify the
strength of a circular rock-paper-scissors effect using the `hyper2`
package, improving on previous results [@west2008] which merely
established its existence.  I use two datasets, a three-player
scoreline and a synthetic dataset referring to wear on a seven-tooth
gear wheel.

# Introduction

@@

# RPS in chess

@@

Above we see a p-value of about 1.2\%, sufficient to reject the null
of zero monster strength^[Note that diagonal entries of `chess_table`
correspond to procyclic wins and off-diagonal entries correspond to
anti-cyclic wins.  This is due to the fact that `chess_table` is of
the same type as the seven-toothed gear wheel dataset considered
below, in which prograde "wins" appear on the diagonal and retrograde
wins appear one place to the right (wrapped if necessary)].

Estimating $M$, the strength of the monster, is also straightforward.
First we consider the case where the probability of a procyclic win is
known to be $p$ (a "procyclic win" is a win that would be expected
under a strong monster).

$$
\operatorname{Prob}(\mathrm{procyclic}\ \mathrm{win})=\frac{\alpha+M}{2\alpha+M}
$$

Here we also have $3\alpha+M=1$, so to estimate the strength of the
monster we would solve

@@

# Profile likelihood

@@

## Analysis of seven-tooth gear with `hyper3`
